---
title: "Agribisnis"
subtitle: "Sistem dan Manajemen"
author: "Tedy Herlambang"
institute: "UPM"
date: " Maret 2022"
output:
  xaringan::moon_reader:
    css: 
      - default
      - css/nhsr.css
      - css/nhsr-fonts.css
    lib_dir: libs
    seal: false
    nature:
      highlightStyle: googlecode
      highlightLines: true
      highlightLanguage: ["r"]
      countIncrementalSlides: false
      ratio: "16:9"
    includes:
      # after_body: [css/insert-logo.html]
---

```{r setup, include = FALSE}
library(knitr)
library(tidyverse)
library(NHSRtheme)
# set default options
opts_chunk$set(echo = FALSE,
               fig.width = 7.252,
               fig.height = 4,
               dpi = 300)

# uncomment the following lines if you want to use the NHS-R theme colours by default
# scale_fill_continuous <- partial(scale_fill_nhs, discrete = FALSE)
# scale_fill_discrete <- partial(scale_fill_nhs, discrete = TRUE)
# scale_colour_continuous <- partial(scale_colour_nhs, discrete = FALSE)
# scale_colour_discrete <- partial(scale_colour_nhs, discrete = TRUE)
```

class: title-slide, left, bottom

# `r rmarkdown::metadata$title`
----
## **`r rmarkdown::metadata$subtitle`**
### `r rmarkdown::metadata$author`
### `r rmarkdown::metadata$date`

```{r share-again, echo=FALSE}
xaringanExtra::use_share_again()
```

```{r style-share-again, echo=FALSE}
xaringanExtra::style_share_again(
  share_buttons = c("twitter", "linkedin", "pocket")
)
```

```{r xaringanExtra, echo=FALSE}
xaringanExtra::use_xaringan_extra(c("tile_view", "animate_css", "tachyons", "scribble"))
```

---

# Tujuan Pembelajaran

Setelah mempelajari bab ini, mahasiswa diharapkan dapat:

* Mendefinisikan dan menjelaskan agribisnis sebagai sistem dan manajemen

* Mendiskusikan asal muasal agribisnis dan perkembangannya

* Memahami perbedaan antara agribisnis dengan bisnis tradisional

* Mengidentifikasi komponen utama agribisnis di jaman sekarang

* Memahami pentingnya model manajemen agribisnis

* Menghargai pertumbuhan peluang di dalam agribisnis


---

# Definisi Agribisnis

* Agribusiness involves the production, distribution, and consumption of food, clothing, and even shelter.

* It includes all economic activity in the food and fiber system: 
  - The input supply industries, 
  - Agricultural production, 
  - Post-harvest, value-added activities: commodity processing, food manufacturing, and food distribution.

* Most people understand and see its parts, but relatively few see the industry in its entirety. 

* In 1957 one ground-breaking book, A Concept of Agribusiness --> The book debuted
the term “agribusiness.”

---

# Agribisnis Sekarang

* Agribusiness firms ranging from large multinational corporations to
emerging food manufacturing and input supply firms.
* A strong demand for more and better employees trained in both management and
agricultural sciences.

---

# Industri-Industri Inti (*The Core Industries*)

* Everyone knows what “business” is --> how about Agribusiness?
* Agribusiness involves:
  1. Inputs—seeds, fertilizer, financing, equipment—that are used in 
 production; 
  2. Intermediate activities— grading, storage, processing, packaging, distribution, pricing, marketing; and 
  3. Final consumption activities—restaurants, groceries.

* The input supply industries, production agriculture, and value-added activities are considered the agribusiness core industries. 
* These core industries use the material and services of certain ancillary industries which are considered agribusiness support industries.
* Farming are part of the core of agribusiness, representing its production process stage. 
* Thus, farming are part of agribusiness. But, they are not the end; they are not even the beginning.

---

# Model Manajemen Agribisnis

* For centuries, agricultural productivity came from the application of two resources: .green[land and labour]. 

* More recently a substantial proportion of increased productivity in agriculture output has come from the use of another resource: **capital**
  - Primarily in the form of technological change including biotechnology, improved animal husbandry, and improvements in machinery and chemicals.

* Future improvements will come from the application of a fourth resource: **management**.

* The industry needs individuals who can harness resources to move a perishable product around the globe --> .green[Agribusiness Management] 
---

# The Chain of Agribusiness—The Value Chain

* The linkages between the agribusiness industry activities: the chain of agribusiness. 
* This chain can be further refined as the agribusiness value chain.
* It begins with raw materials and commodities of relatively low value; 
* Then each step in the chain modifies the material or commodity and adds value to the
product derived from its step.

---

## Contoh Rantai Pasok Buah
- Growers harvest raw fruits or vegetables that are transported in bulk to processors or wholesalers, 
- Then they are modified or repackaged and transported to grocery chains in processed form or in smaller batches, 
- Finally at local grocers who display and sell the products in quantities more suitable for consumers to purchase.
---

# Another way of looking at the term agribusiness

* Agribusiness involves all the value chain activities usually associated with a business or industry with the additional provision that most of the materials, commodities, and end products are **perishable** (ringkih, gampang rusak, cepat busuk). 
* The perishable nature of a product, along with variability in such factors as color, texture, size, and shape, is the distinguishing characteristic of agribusiness products and industries from other businesses. 

> This aspect of the agribusiness industry is also a key factor in all business decisions. 

---

# Agribisnis di sekitar kita
* For this reason the story of Ayam Geprek, or Bebek Goreng or Kios Buah is an agribusiness story. 
* Bagaimana dengan Warteg, Warung Padang, Warung Nasi Pecel dan Rawon?
* Its entire business system involved a highly perishable commodity: daging ayam, beras, bumbu-bumbu, minyak goreng, dll.

---

## .red[Diskusi]

* Berikan contoh perusahaan, usaha, kegiatan yang bisa dikategorikan sebagai agribisnis disekitar Anda! 
* Bagaimana dengan peluang usaha di bidang Agribisnis di sekitar Anda?

---
class: inverse, right, bottom
background-image: url(assets/simpleflower.svg)
background-size: 260px
background-position: 5% 95%

### Matur nuwun, find me at...
<hr width="45%" align="right" size="0.3" color="orange"></hr>

<a href="http://twitter.com/t_hlb">
.white[`r icons::fontawesome("twitter")` @t_hlb]
</a>

<a href="http://github.com/bangtedy">
.white[`r icons::fontawesome("github")` @bangtedy]
</a>
